---
title: "Course Assignment"
author: "Moritz Körber"
test: "`r Sys.setlocale(locale = 'English')`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First step: Load required packages:

```{r message=FALSE, warning=FALSE}
library("dplyr")
library("caret")
library("mlr")
library("visdat")
library("data.table")
```

Load data and set seed:

```{r}
df <- fread("1_data/pml-training.csv", na.strings = c("NA", "NaN", "", "#DIV/0!"), drop = 1)

set.seed(31)
```

# 1. Inspect target variable
First, I am looking at the target variable "classe":

```{r}
glimpse(df$classe)
```

Since it is a categorical variable, it is a classification problem. Hence, an algorithm like logistic regression, random forests, or a similar algorithm is suitable. Let's check if the classes are balanced:

```{r}
ggplot(df, aes(x = classe)) +
  geom_bar(stat = "count") +
  theme_classic()

count(df, classe)
```

# 2. Clean the data set
Remove obvious non-predictive variables:
```{r}
df %>% select(-c(raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, user_name)) -> df
```

Remove zero variance/near zero variance predictors:
```{r}
nzv <- nearZeroVar(df)
df <- df[ , -..nzv]

df <- removeConstantFeatures(as.data.frame(df), perc = .02)
```

Remove predictors with mostly NAs:
```{r}
df %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.)) / length(.))) -> p

# check these variables
vis_miss(df[which(p > 0.975)],
  sort_miss = TRUE, warn_large_data = F
)

# remove them if sensible
df[which(p > 0.975)] <- NULL
```

Remove highly correlated predictors:
```{r}
nums <- select_if(df, is.numeric)
descrCor <- cor(nums)

highCorr <- sum(na.omit(abs(descrCor[upper.tri(descrCor)])) >= .98)

na.omit(descrCor[upper.tri(descrCor)])[which(na.omit(abs(descrCor[upper.tri(descrCor)])) >= .98, arr.ind = TRUE)]

which(na.omit(abs(descrCor)) >= .98 & na.omit(abs(descrCor)) < 1, arr.ind = TRUE)

findCorrelation(na.omit(descrCor), cutoff = .98, verbose = T, exact = T, names = T)
```

There are two variables with high correlation. I will leave them in the dataset for this time.

Find linear combinations:
```{r}
findLinearCombos(nums)
```

Visualize the data:
```{r}
vis_dat(df, warn_large_data = F)
```

Are there mixed data types in one variable?
```{r}
vis_guess(df)
```

Plot the features:
```{r}
nums <- unlist(lapply(df, is.numeric))
featurePlot(x = df[nums], y = df$classe, plot = "strip")
```

Save the cleaned data:
```{r eval = FALSE}
saveRDS(df, "1_data/cleaned_data.rds")
```

# 3. Training
The training is run on an Amazon AWS EC2 t2.2xlarge instance. More details on the environment:
```{r echo = FALSE}
readRDS("sessionInfo.rds")
```

## Task
```{r}
task <- makeClassifTask(id = "fitness.tracker", data = df, target = "classe")
```

## Resampling
I chose to use a nested cross validation strategy with a 5-fold inner cross validation and a 3-fold outer cross validation. The evaluation of the tuned learners is performed by 5-fold cross validation. 
```{r}
rdesc.inner <- makeResampleDesc("CV", iters = 5)
```

The best parameter combination is then evaluated against the remaining fold of the 3-fold outer cross validation.
```{r}
rdesc.outer <- makeResampleDesc(method = "CV", iters = 3)
resample.instance.outer <- makeResampleInstance(desc = rdesc.outer, task = task)
```

## Measures
The mean misclassification error is one of the most important metrics in classification problems.
```{r}
measures <- list(mmce)
```

## Learners
I compare the performance of three different learners. Each learner's hyperparameters are tuned in this evaluation process. For preprocessing, I center and scale the features.

### Random forest
```{r eval = FALSE}
lrn.rndforest <- makePreprocWrapperCaret("classif.randomForest", ppc.center = T, ppc.scale = T)

ps.rndforest <- makeParamSet(
  makeIntegerParam("ntree", lower = 100, upper = 1000),
  makeIntegerParam("mtry", lower = 5, upper = 20)
)

tune.ctrl.rndforest <- makeTuneControlRandom(maxit = 30)

tuned.lrn.rndforest <- makeTuneWrapper(lrn.rndforest,
  par.set = ps.rndforest,
  resampling = rdesc.inner,
  control = tune.ctrl.rndforest
)
```

### XGBoost
```{r}
lrn.xgboost <- makePreprocWrapperCaret("classif.xgboost", ppc.center = T, ppc.scale = T)

ps.xgboost <- makeParamSet(
  makeNumericParam("eta", lower = 0, upper = 0.5),
  makeNumericParam("colsample_bytree", lower = 0.5, upper = 0.9),
  makeNumericParam("gamma", lower = 0, upper = 2),
  makeIntegerParam("max_depth", lower = 4, upper = 10),
  makeIntegerParam("nrounds", lower = 500, upper = 1500)
)

tune.ctrl.xgboost <- makeTuneControlRandom(maxit = 30)

tuned.lrn.xgboost <- makeTuneWrapper(lrn.xgboost,
  par.set = ps.xgboost,
  resampling = rdesc.inner,
  control = tune.ctrl.xgboost
)
```

### Ranger
```{r eval = FALSE}
lrn.ranger <- makePreprocWrapperCaret("classif.ranger", ppc.center = T, ppc.scale = T)
```

## Benchmark
Compare the leaners' performance:
```{r eval = FALSE}
bm <- benchmark(
  learners = list(
    tuned.lrn.rndforest,
    tuned.lrn.xgboost,
    lrn.ranger
  ),
  tasks = task,
  resamplings = resample.instance.outer,
  measures = measures
)
```

```{r echo = FALSE}
readRDS("bm.rds")
```

```{r eval = FALSE}
plotBMRBoxplots(bm)
```

```{r echo=FALSE}
readRDS("bmplot.rds")
```

XGBoost seems to be the best learner for this problem. Thus, I choose it to train my final model on the whole data set.

# 4. Train final model
```{r}
model <- mlr::train(learner = tuned.lrn.xgboost, task = task)

save(model, file = "model.rds")
```

# 5. Predict the test data
## Load and prepare test data
```{r}
testing <- read.csv("1_data/pml-testing.csv", na.strings = c("NA", "NaN", "", "#DIV/0!"), row.names = 1)

# make sure that they have the same columns (except the target)
df %>%
  select(-classe) %>%
  colnames() -> vars

testing <- testing[vars]
```

## Prediction
Lastly, I predict the test data.
```{r}
pred <- predict(model, testing)

pred
```
